---
title: "SOTF Child's River Analysis"
author: "Ben Egan & Hannah Schofield"
output:
  word_document: default
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
  html_document:
    df_print: paged
subtitle: https://github.com/BenjaminEgan1/EganSchofield_ENV872_FinalProject.git
geometry: margin=2.54cm
fontsize: 12pt
mainfont: Times New Roman
---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory

#setwd("C:/Users/hannahschofield/ENV872/ENV872_FinalProj/EganSchofield_ENV872_FinalProject")


# Load your packages
library(lubridate)
library(ggplot2)
library(dplyr)
library(sf)
library(mapview)
library(openxlsx)

# Set your ggplot theme

mytheme <- theme_classic(base_size = 10) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

# Load your datasets
#getwd()
SOTFRaw <- read.xlsx("/Users/hannahschofield/ENV872/ENV872_FinalProj/EganSchofield_ENV872_FinalProject/Data/Raw/SOTF_Data (1).xlsx", sheet = 3)


```


# Rationale and Research Questions
**I aim to research the data in question**


\newpage

# Dataset Information
```{r information}
head(SOTFRaw)
dim(SOTFRaw)
```



\newpage

# Exploratory Analysis 
```{r Exploratory}
SOTFRaw$Sample.Date <- as.Date(SOTFRaw$Sample.Date, origin, format = "%m/%d/%Y")

class(SOTFRaw$Sample.Date)
colnames(SOTFRaw)

SOTF_Childs <- SOTFRaw %>% filter(map_code %in% c("MA_CH09", "MA_CH08", "MA_CH07", "MA_CH06", "MA_CH05", "MA_CH04", "MA_CH03", "MA_CH02", "MA_CH01")) %>% select(Sample.ID, SiteName, Sample.Date, map_code, Latitude, Longitude, 'State/Region', `NO3-N_µgL`, `NH4-N_µgL`, `TDN_µgL`, `PO4-P_µgL`, `SiO2_mgL`, `DOC_mgL`, `Water.temperature.(°F)`)

na.omit(SOTF_Childs)

Childs_Plot_NO3 <- ggplot(SOTF_Childs, aes(x = Sample.Date, y = `NO3-N_µgL`, color = SiteName )) + geom_point()  + geom_smooth(se = FALSE) 

Childs_Plot_TDN <- ggplot(SOTF_Childs, aes(x = Sample.Date, y = TDN_µgL, color = SiteName)) + geom_point()  + geom_smooth(se = FALSE)

Childs_Plot_Temps <- ggplot(SOTF_Childs, aes(x = Sample.Date, y = `Water.temperature.(°F)`, color = SiteName )) + geom_point()

print(Childs_Plot_NO3)
print(Childs_Plot_TDN)
print(Childs_Plot_Temps)
```
```{r correlationAnalysis}
#install.packages("ggpubr")
library(ggpubr)
library(GGally)

#NO3-N_ugl correlation
cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`NH4-N_µgL`)

cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`TDN_µgL`)
ggscatter(SOTF_Childs, x = "NO3-N_µgL", y = "TDN_µgL", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "NO3 µgL", ylab = "TDN µgL")
#NO3-N_µgL and TDN_µgL are highly correlated with a correlation coefficient of 0.96. 

cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`PO4-P_µgL`)
cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`SiO2_mgL`)
cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`DOC_mgL`)
cor.test(SOTF_Childs$`NO3-N_µgL`,SOTF_Childs$`Water.temperature.(°F)`)
#No other variables are correlated

#NO4 Correlation
cor.test(SOTF_Childs$`NH4-N_µgL`,SOTF_Childs$`TDN_µgL`)
cor.test(SOTF_Childs$`NH4-N_µgL`,SOTF_Childs$`SiO2_mgL`)
cor.test(SOTF_Childs$`NH4-N_µgL`,SOTF_Childs$`DOC_mgL`)
cor.test(SOTF_Childs$`NH4-N_µgL`,SOTF_Childs$`Water.temperature.(°F)`)
#No correlated variables 

#TDN Correlations 
cor.test(SOTF_Childs$`TDN_µgL`,SOTF_Childs$`SiO2_mgL`)

cor.test(SOTF_Childs$`TDN_µgL`,SOTF_Childs$`DOC_mgL`)
ggscatter(SOTF_Childs, x = "TDN_µgL", y = "DOC_mgL", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "TDN Concentration", ylab = "DOC Concentration")
#Mild negative correlation between TDN and DOC with a correlation coeff of -.33. 

cor.test(SOTF_Childs$`TDN_µgL`,SOTF_Childs$`Water.temperature.(°F)`)


#SiO2 Correlation 
cor.test(SOTF_Childs$`SiO2_mgL`,SOTF_Childs$`DOC_mgL`)
cor.test(SOTF_Childs$`SiO2_mgL`,SOTF_Childs$`Water.temperature.(°F)`)
#no correlation


#DOC Correlation 
cor.test(SOTF_Childs$`DOC_mgL`,SOTF_Childs$`Water.temperature.(°F)`)


```

```{r spatial Analysis}
SOTF.sf <- SOTF_Childs %>% 
  st_as_sf(coords = c('Longitude','Latitude'), crs=4326)

class(SOTF.sf)
mapview(SOTF.sf)


```



\newpage

# Analysis




## Question 1: Does concentration of nutrient fluctuate as you move down stream

## Question 2: What Nutrients are correlated to eachother?




\newpage

# Summary and Conclusions
  


\newpage

# References
<add references here if relevant, otherwise delete this section> 
